What We Eval in the Shadows

### 2021-04-16 paper/main.tex
Some love for Section 6

\todo{Illustrate things that we have discussed in Section 5 - examples}
\todo{Things to discuss: model.frame, match.call, subset as example of NSE, substitute, parse, plots}

% We present here the main uses of \eval we have witnessed in R. Sometimes, they are design patterns, sometimes, \eval could be replaced by more specific functions. For each use of \eval, we discuss whether it is necessary to use.

% \subsection{Variable lookup}

% \eval is used to look-up the value of a variable in a given environment, different from the current scope so the \c{envir} argument of \eval is explicitly specified.

% \subsection{Metaprogramming}

% \mypara{\c{substitute}}

% \mypara{\c{match.call}}

% \todo{Talk about model.frame}

% \subsection{Logging}

% \subsection{Code transform}

% % Derivation of symbolic expressions

% \subsection{Plotting}

@@

\subsubsection{Complex function calls.} It corresponds to minimized expressions \c{F(X)} and \c{F(F(X))}, \ie function calls with variables, possibly with nested function calls. Package \emph{huxtable} is used to create custom tables and export them to various formats. To create functions to customize each of the borders of the table, it injects the indices \c{i} and \c{j} of the cells at the border and then evaluate them:
\begin{lstlisting}
xxx_border_huxtable <- function(lr_tb, i, j, border_prop) {
	function (ht) {
		hickness <- mlist$thickness[eval(i), eval(j), drop = FALSE]
		...
	}
}
left_border.huxtable <- xxx_border_huxtable("lr_borders",
i = quote(seq_len(nrow(ht))), j = quote(-last_col))
\end{lstlisting}

@@

\subsubsection{Domain Specific Language.} R expressions are built and transformed symbolically and then evalled to a concrete value. For instance,   base R packages \emph{stats} provides two operators for \emph{symbolic differentiation}, \c{D} and \c{deriv}. They support arithmetic operations and functions on real numbers such as \c{sin}.
In \emph{MCMCglmm} (Monte Carl Markov Chain Generalised Linear Mixed Models), a \c{Dtensor} object is created, which contains expressions derived twice (with \c{DD}) and then typically in a list or a data frame.
\begin{lstlisting}
 expr<-expression(beta_1 + time*beta_2+u)
 ...
 mu = data.frame(beta_1=0.5, beta_2=1, time=3, u=2.3)
 ...
 D[i]<-eval(DD(expr, name[unlist(comb.pos[i,])]), mu)
\end{lstlisting}
% See https://github.com/cran/MCMCglmm/blob/cfac9e67ec73a3db2142826faae35e5d2318da31/man/evalDtensor.Rd

>>>>

@@

\subsubsection{Complex function calls.} It corresponds to minimized expressions \c{F(X)} and \c{F(F(X))}, \ie function calls with variables, possibly with nested function calls. Package \emph{huxtable} is used to create custom tables and export them to various formats. To create functions to customize each of the borders of the table, it injects the indices \c{i} and \c{j} of the cells at the border and then evaluate them:
\begin{lstlisting}
xxx_border_huxtable <- function(lr_tb, i, j, border_prop) {
	function (ht) {
		hickness <- mlist$thickness[eval(i), eval(j), drop = FALSE]
	}
}
left_border.huxtable <- xxx_border_huxtable("lr_borders",
i = quote(seq_len(nrow(ht))), j = quote(-last_col))
\end{lstlisting}

@@

\subsubsection{Domain Specific Language.} R expressions are built and transformed symbolically and then evalled to a concrete value. For instance,   base R packages \emph{stats} provides two operators for \emph{symbolic differentiation}, \c{D} and \c{deriv}. They support arithmetic operations and functions on real numbers such as \c{sin}.
In \emph{MCMCglmm} (Monte Carl Markov Chain Generalised Linear Mixed Models), a \c{Dtensor} object is created, which contains expressions derived twice (with \c{DD}) and then typically in a list or a data frame.
\begin{lstlisting}
 expr<-expression(beta_1 + time*beta_2+u)
 mu = data.frame(beta_1=0.5, beta_2=1, time=3, u=2.3)
 D[i]<-eval(DD(expr, name[unlist(comb.pos[i,])]), mu)
\end{lstlisting}
% See https://github.com/cran/MCMCglmm/blob/cfac9e67ec73a3db2142826faae35e5d2318da31/man/evalDtensor.Rd
